%-*- Mode:LaTeX; -*-      
\documentclass[11pt]{article}
\usepackage{geometry}		% Force narrower margins
\geometry{letterpaper}
\geometry{lmargin=1.0in, tmargin=1.0in, rmargin=1.0in, bmargin=0.6in}
\geometry{headsep=0in, footskip=0.4in}

\setlength{\parskip}{.1in}  % removed space between paragraphs
\setlength{\parindent}{0in}

\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{epstopdf}
\newcommand{\ra}{$\rightarrow$~}
\newcommand{\dt}{$\circ$~}
\newcommand{\adj}{\texttt{/ADJ }}
\newcommand{\adv}{\texttt{/ADV }}
\newcommand{\art}{\texttt{/ART }}
\newcommand{\conj}{\texttt{/CONJ }}
\newcommand{\ger}{\texttt{/GER }}
\newcommand{\infi}{\texttt{/INF }}
\newcommand{\mod}{\texttt{/MOD }}
\newcommand{\noun}{\texttt{/NOUN }}
\newcommand{\particle}{\texttt{/PART }}
\newcommand{\prep}{\texttt{/PREP }}
\newcommand{\perpro}{\texttt{/PERPRO }}
\newcommand{\relpro}{\texttt{/RELPRO }}
\newcommand{\ver}{\texttt{/VERB }}
\newcommand{\sub}{\texttt{/Subject }}
\newcommand{\dobj}{\texttt{/Direct Object }}
\newcommand{\iobj}{\texttt{/Indirect Object }}
\newcommand{\act}{\texttt{ACTIVE voice}}
\newcommand{\pas}{\texttt{PASSIVE voice}}

\begin{document}

\large
\begin{center}
{\bf CS-5340/6340, Written Assignment \#1} \\
{\bf DUE: Tuesday, September 4, 2018 by 11:59pm}
\end{center}
\normalsize

\begin{enumerate}  

%% ===============================================================
% QUESTION #1 : Parts of Speech    
%% =============================================================

\item (33 pts) 

\begin{enumerate}

\item John\noun might\mod go\ver to\prep school\noun to\infi learn\ver welding\ger.
\vspace*{.2in}


\item Fires\noun broke\ver out\particle near\prep Moab\noun without\prep warning\ger.
\vspace*{.2in}


\item Kate\noun, who\relpro is\ver brilliant\adj, builds\ver robots\noun. 
\vspace*{.2in}


\item Susan\noun loves\ver her\perpro house\noun but\conj may not\mod stay\ver there\adv.
\vspace*{.2in}


\item Sleeping\ger bags\noun can\mod keep\ver people\noun from\prep freezing\ger on\prep camping\ger trips\noun.
\vspace*{.2in}


\item To\infi quit\ver smoking\ger, Joe\noun never\adv buys\ver cigarettes\noun, which\relpro he\perpro craves\ver.
\vspace*{.2in}


\item  Tom\noun makes\ver up\particle stories\noun about\prep imaginary\adj monsters\noun.
\vspace*{.2in}


\item She\perpro delicately\adv brought up\particle a\art sensitive\adj topic\noun to\prep Bill\noun.
\vspace*{.2in}


\end{enumerate}


\newpage
%% ===============================================================
% QUESTION #2: Active/Passive Voice
%% ================================================================

\item (20 pts) 

\begin{enumerate}

\item The bird quickly built an amazing nest for its young.\\
\act
\vspace*{.5in}

\item George has been having serious problems with his back.\\
\act
\vspace*{.5in}

\item Mary does not want help with her car.\\
\act
\vspace*{.5in}

\item The window of the Toyota Prius was broken during a hail storm.\\
\pas
\vspace*{.5in}

\item Too much money has been spent on unnecessary trips to Europe.\\
\pas
\vspace*{.5in}

\item He could barely understand the writing on the chalkboard.\\
\act
\vspace*{.5in}

\item Both girls from Utah were chosen for the summer program.\\
\pas
\vspace*{.5in}

\item John was told about the election results by his neighbor.\\
\pas
\vspace*{.5in}

\item Tina received several awards for her athletic skills.\\
\act
\vspace*{.5in}

\item The boy had seen many bears near the family's cabin.\\
\act
\vspace*{.5in}

\end{enumerate}



\newpage
%% ===============================================================
% QUESTION #3: Syntactic Roles
%% ================================================================

\item (20 pts) 

\begin{enumerate}

\item \underline{His grandmother}\sub left \underline{her son}\iobj \underline{many valuable jewels}\dobj.
\vspace*{.5in}

\item \underline{The cat}\sub was chased by a barking dog.
\vspace*{.5in}

\item \underline{Tom Brady}\sub passed \underline{the football}\dobj to his wide receiver.
\vspace*{.5in}

\item Don't forget \underline{your car keys}\dobj again.
\vspace*{.5in}

\item \underline{George}\sub mailed \underline{his daughter}\iobj \underline{a letter}\dobj at her summer camp. 
\vspace*{.5in}

\item \underline{IBM}\sub awarded \underline{a bonus}\dobj to John for his inventions.
\vspace*{.5in}

\item \underline{You}\sub should pick \underline{your favorite movie}\dobj!
\vspace*{.5in}

\item \underline{The old man}\sub promised \underline{the boy}\iobj \underline{a lawn mowing job}\dobj in summer.
\vspace*{.5in}

\item \underline{Susan}\sub was extremely happy about \underline{the news}\dobj.
\vspace*{.5in}

\item \underline{The woman with very long hair}\sub was blocking \underline{the view}\dobj in the movie theater.
\vspace*{.5in}


\end{enumerate}


\newpage
%% ===============================================================
% QUESTION #4: Parse Trees
%% ================================================================

\item (12 pts) List all of the context-free grammar rules that are depicted in the parse trees above.  You only need to list grammar rules for the non-terminal symbols S, NP, VP, and PP. You do \underline{not} need to list rules for the non-terminal symbols associated with part-of-speech tag assignments (i.e., noun, verb, etc.).

\vspace*{.2in}
$$S \rightarrow NP \  VP$$
$$S \rightarrow NP \ S \  VP$$
$$S \rightarrow VP \  PP$$
\vspace*{.2in}
$$NP \rightarrow noun$$
$$NP \rightarrow art \ noun$$
$$NP \rightarrow noun \ PP$$
$$NP \rightarrow art \ noun \ PP$$
$$NP \rightarrow perpro \ noun$$
\vspace*{.2in}
$$VP \rightarrow verb$$
$$VP \rightarrow verb \ NP$$
$$VP \rightarrow adv \ verb \ NP \ PP$$
\vspace*{.2in}
$$PP \rightarrow prep \ NP$$

\newpage
%% ================================================================
% QUESTION #5: Subcategorization Frames
%% ================================================================

\item (15 pts)   

\begin{enumerate}

\item {\it drive}\\
\textbf{NP} I am driving a truck.\\
\textbf{PP} I drive to school everyday.

\item {\it laugh}\\
\textbf{None}

\item {\it wear}\\
\textbf{NP} She is wearing a red hat.

\item {\it hope}\\
\textbf{that S} I hope that she can pass the exam.

\item {\it point}\\
\textbf{NP} He pointed the way to the stadium.\\
\textbf{PP} The arrow is pointing to the North.

\end{enumerate}


\newpage
\underline{\textbf{Question \#6 is for CS-6340 students ONLY!}}  \\

\item (12 pts) 

\begin{enumerate}

\item G1 
$$\texttt{art}?\texttt{adj}*\texttt{noun}+$$
\vspace*{.2in}

\item G2
$$\texttt{art}?\texttt{adj}*\texttt{noun}+$$
\vspace*{.2in}

\item G3
$$\texttt{art}?\texttt{adj}?\texttt{noun}+$$
\vspace*{.2in}

\item G4
$$\texttt{art}?\texttt{adj}*\texttt{noun}+$$
\vspace*{.2in}

\end{enumerate}


\end{enumerate}  % END OF WRITTEN QUESTIONS

\newpage
\hspace*{1.5in}  {\bf ELECTRONIC SUBMISSION INSTRUCTIONS} \\

You should submit the answers to this assignment {\bf in pdf format}
on our course's CANVAS site by 11:59pm on Tuesday, September 4.

\end{document}


